---
title: "HW_2"
output:
  pdf_document: default
  html_document: default
date: "2025-04-14"
---
# Question 2

### 1

```{r}
corona_df <- read.csv("data_corona_world.csv")
head(corona_df)
```

### 2

```{r}
filterd_df <- subset(corona_df, Country %in% c("Italy", "China"))
head(filterd_df)
```

### 3


```{r}
str(filterd_df)
```

### 4

```{r}
china_data <- subset(filterd_df, Country == 'China',
                     select = c(Confirmed.cases, Confirmed.fatalities))
china_data <- colSums(china_data)
china_survivors <- china_data["Confirmed.cases"] - china_data["Confirmed.fatalities"]

italy_data <- subset(filterd_df, Country == 'Italy',
                     select = c(Confirmed.cases, Confirmed.fatalities))
italy_data <- colSums(italy_data)
italy_survivors <- italy_data["Confirmed.cases"] - italy_data["Confirmed.fatalities"]

contingency_table <- rbind(
  c(china_survivors, china_data["Confirmed.fatalities"]),
  c(italy_survivors, italy_data["Confirmed.fatalities"])
)

rownames(contingency_table) <- c("China","Italy")
colnames(contingency_table) <- c("Survivor","Death")

print(contingency_table)

mosaicplot(contingency_table, color = TRUE,
           main = "Contingency Table: China vs. Italy")
```

### 5

```{r}
china_dp <- (contingency_table["China", "Death"] / sum(contingency_table["China", ])) * 100
print(china_dp)

italy_dp <- (contingency_table["Italy", "Death"] / sum(contingency_table["Italy", ])) * 100
print(italy_dp)
```

It seems that China handled the situation better, since the death rate was lower compared to Italy.

### 6

```{r}
death_rate_df <- data.frame(
  Country = c("China", "Italy"),
  DeathRate = c(china_dp, italy_dp)
)

library(lattice)
barchart(DeathRate ~ Country, 
         data = death_rate_df, 
         main = "Death Rate by Country", 
         xlab = "Country", 
         ylab = "Death Rate (%)",
         ylim = c(0,5)
         )
```

Based on the graph, being in Italy appears to be more risky.

```{r}
df_age_china <- subset(filterd_df, Country == 'China',
                 select = c(Age.group, Confirmed.cases, Confirmed.fatalities))

df_age_china$DeathRatio <- (df_age_china$Confirmed.fatalities / (df_age_china$Confirmed.cases + df_age_china$Confirmed.fatalities)) * 100
df_age_china$Country <- "China"


df_age_italy <- subset(filterd_df, Country == 'Italy',
                       select = c(Age.group, Confirmed.cases, Confirmed.fatalities))

df_age_italy$DeathRatio <- (df_age_italy$Confirmed.fatalities / (df_age_italy$Confirmed.cases + df_age_italy$Confirmed.fatalities)) * 100
df_age_italy$Country <- "Italy"


death_rate_age <- rbind(df_age_china, df_age_italy)

death_rate_age$Age.group <- factor(
  death_rate_age$Age.group,
  levels = c("0-9", " 10-19", "20-29", "30-39", "40-49", "50-59", "60-69", "70-79", "80+"),
  ordered = TRUE
)


barchart(DeathRatio ~ Age.group,
         group = Country,
         data = death_rate_age,
         auto.key = list(space = "right"),
         main = "Death Rate by Country & Age",
         xlab = "Age Group",
         ylab = "Death Rate (%)"
         )
```

This is an example of Simpson's paradox, because when we look at each age group separately, China's death rate was higher than Italy's.

# Question 3

```{r}
df_titanic <- read.csv("titanic.csv")
head(df_titanic)
```

### a

```{r}
logic_vec <- !colnames(df_titanic) %in% c("Cabin", "Ticket")

df_titanic <- df_titanic[,logic_vec]
head(df_titanic)
```

### b

```{r}
completed_rows <- complete.cases(df_titanic)

df_titanic <- df_titanic[completed_rows, ]
head(df_titanic)
```

### c

```{r}
fare_summary <- aggregate(Fare ~ Pclass, data = df_titanic, FUN = summary)
print(fare_summary)
```

According to the summary, the median and mean for each class show that Class 1 is the most expensive.

```{r}
boxplot(Fare ~ Pclass, data = df_titanic,
        main = "Fare by Passenger Class",
        xlab = "Pclass", ylab = "Fare",
        col = c("lightblue", "lightgreen", "salmon"),
        ylim = c(0,600))
```

We should use a transformation, since there are outliers that make the graph harder to read. We chose a log transformation.

```{r}
df_titanic$logFare <- log(df_titanic$Fare + 1)

boxplot(logFare ~ Pclass, data = df_titanic,
        main = "Fare by Passenger Class",
        xlab = "Pclass", ylab = "Fare (log)",
        col = c("lightblue", "lightgreen", "salmon"),
        ylim = c(0,10)
        )
```



```{r}
variance_by_class <- aggregate(Fare ~ Pclass, data = df_titanic, FUN = var)
print(variance_by_class)
```

Class 1 shows higher fare variance, meaning there was a wider range in ticket prices.

### d

```{r}
table_survival <- table(df_titanic$Survived, df_titanic$Pclass)
rownames(table_survival) <- c("Died", "Survived")
colnames(table_survival) <- c("Class 1", "Class 2", "Class 3")
mosaicplot(table_survival,
           color = TRUE,
           main = "Survival by Passenger Class"
           )
```

### d.1

The marginal totals are calculated as

```{r}
colSums(table_survival)
rowSums(table_survival)
```

### d.2

Expected frequency

```{r}
row_total <- c(424, 290) # Died, Survived
col_total <- c(186, 173, 355) # Class 1,2,3
total <- sum(row_total)

expected <- matrix(0, nrow = 2, ncol = 3)

for (i in 1:2) {
  for (j in 1:3) {
    expected[i,j] <- (row_total[i] * col_total[j]) / total
  }
}

rownames(expected) <- c("Died", "Survived")
colnames(expected) <- c("Class 1", "Class 2", "Class 3")

print(expected)
```

### d.3

The X² statistic is

```{r}
x <- 0
for (i in 1:2) {
  for (j in 1:3) {
    x <- x + ((table_survival[i,j] - expected[i,j]) ^ 2) / expected[i,j]
  }
}

print(x)
```

### d.4

```{r}

```

### e

```{r}
survival_ratio <- table_survival[2, ] / (table_survival[1, ] + table_survival[2, ])
print(survival_ratio)
```

The proportions clearly show that Class 1 had the highest chance of survival

### f

```{r}
boxplot(Age ~ Survived,
        names = c("Died", "Survived"),
        data = df_titanic,
        ylab = "Age",
        xlab = "Survival Status",
        col = c("salmon", "lightgreen"),
        main = "Age Distribution by Survival"
        )
```

Based on the graph, there doesn’t seem to be a connection between survival and age, since the medians look very similar and the Q1 and Q3 are at about the same level.

```{r}
boxplot(Age ~ Survived + Pclass, 
        data = df_titanic,
        names = c("Died-1st", "Survived-1st", 
                  "Died-2nd", "Survived-2nd", 
                  "Died-3rd", "Survived-3rd"),
        col = c("salmon", "lightgreen"),
        xlab = "Survival Status by Class",
        ylab = "Age",
        main = "Age Distribution by Survival and Class"
        )
```

However, when separating by class, it became apparent that younger passengers were more likely to survive within each class.
