---
title: "HW3"
output:
  pdf_document: default
  html_document: default
date: "2025-04-25"
---
# Question 3

```{r}
library(MASS)
dat <- Animals
head(dat, 2)
```

### a

```{r}
plot(dat$brain ~ dat$body,
     xlab = "Body weight (kg)",
     ylab = "Brain wight (g)",
     main = "Brain weight (g) vs Body weight (kg)")
```

### b

```{r}
par(mfrow = c(1, 2))

hist(dat$brain,
     main = "Brain",
     xlab = "brain",
     breaks = 25)

hist(dat$body,
     main = "Body",
     xlab = "body",
     breaks = 25)

par(mfrow = c(1, 1))
```


It seems preferable to apply a log transformation.


```{r}
par(mfrow = c(1, 2))
log_brain = log(dat$brain + 1)
log_body = log(dat$body + 1)

hist(log_brain,
     main = "log(Brain + 1)",
     xlab = "brain")

hist(log_body,
     main = "log(Body + 1)",
     xlab = "body")

par(mfrow = c(1, 1))
```

### c

```{r}
log_body = log(dat$body + 1)
log_brain = log(dat$brain + 1)

plot(log_brain ~ log_body,
     xlab = "log(Body + 1)",
     ylab = "log(Brain + 1)",
     main = "log(Brain [g] + 1) vs log(Body [kg] + 1)")
```


The plot after the log transformation appears approximately linear.


### d

```{r}
log_lm = lm(log_brain ~ log_body)
coef(log_lm)
```


The regression coefficient for log_body is 0.489, indicating a positive linear relationship between log_body and log_brain.


### e

```{r}
par(mfrow = c(1, 2))

plot(log_lm$residuals ~ log_body,
     xlab = "log(Body weight + 1) (kg)",
     ylab = "Residuals",
     main = "Residuals vs log(Body weight + 1)",
     cex.main = 0.8)
abline(h = 0, col = "red")

raw_lm = lm(dat$brain ~ dat$body)
            
plot(raw_lm$residuals ~ dat$body,
     xlab = "Body weight (kg)",
     ylab = "Residuals",
     main = "Residuals vs Body weight",
     cex.main = 0.8)
abline(h = 0, col = "red")

par(mfrow = c(1, 1))
```


The transformed model fits, while the untransformed model does not, due to residual concentration below zero.


### f

```{r}
plot(log_brain ~ log_body,
     xlab = "log(Body weight + 1) (kg)",
     ylab = "log(Brain weight + 1) (g)",
     main = "log(Brain + 1) vs log(Body + 1) with Regression Line")
abline(log_lm, col = "blue")
```









